A Recipe for Scavenging - the natural history of a behaviour

### 2016-03-29 natural history of scavenging.tex
snakes as detoxifiers

%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."

Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
Many extant birds exist as facultative scavengers; storks, eagles, corvids, are all known to take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still (approximately twice the basal metabolic rate) \citep{hedenstrom1993migration,spivey2014analysing}.
The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
In the former case we have the best known scavengers on Earth. 
Vultures consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
Avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
However, vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
Certain clades of these animals could reach enormous sizes \citep[e.g. Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, look to have engaged in soaring flight \citep{witton2010size}.
%TG: this paragraph's a killer! Love it!

@@

Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
Extant aquatic snakes are deemed as having the most suitable physiology for scavenging.
A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.

@@

Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
This is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.

%TG: or "could"? This is hypothetical behaviour stuff! Maybe past hyenas where exclusive sushi eaters!
%TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.
The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.

@@

Avian scavengers have to compete with microorganisms and so have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses \textit{Diomedea exulans} had an average pH of 1.5, which enables them to consume fisheries discards. 
%This led the authors to term them 'vultures of the seas'.

%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore probably lack of adaptations for digesting  meat.
While carnivorous bats are mainly found in the microbats which are insectivorous.
% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds! Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate! And that's much better than birds where you're stuck with just some fragments of long bones.

>>>>

%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
But it is in the air that we find scavengers par excellence. 
Flight is a cheaper means of locomotion than walking or running \citep{tucker1975energetic}. 
We know that many extant birds exist as facultative scavengers; storks, eagles, corvids, all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still (approximately twice the basal metabolic rate) \citep{hedenstrom1993migration,spivey2014analysing}.
The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
In the former case we have the best known scavengers on Earth. 
Vultures consist of two convergent groups, from the old and the new world and represent the only example of obligate vertebrate scavengers today.
The families from which modern vultures arose, the Accipitridae and Cathartidae, appear during the Palaeocene \citep[66 - 56 Mya; ][]{Jetz2012, Jarvis2014}.
Avian flight is far older than this and originates in the Late Jurassic (163.5-145 Mya), conincident with the fossils of \textit{Archaeopteryx lithographica} so many of these benefits would have been realised from that point on for carnivorous birds.
However, vertebrate flight is much older still where pterosaurs predate bird origins by a considerable margin in the Late Triassic (235-201.3 Mya).
Scavenging in this diverse group has been hypothesied many times before \citep{witton2008reappraisal}.
Certain clades of these animals could reach enormous sizes \citep[e.g. Azhdarchids with wingspans of 11 metres; ][]{witton2010size} and, notably, look to have engaged in soaring flight \citep{witton2010size}.
%TG: this paragraph's a killer! Love it!

@@

Depending on the species, a carcass in water either floats or descends to the sea floor \citep{Whitehead415}.
In the latter low-light environment, visual detection distances are far lower (< 100 m) than they would be in the air.
As such, animals detect resources through chemo- and mechanoreception more so than through vision \citep{ruxton2004energetic}.
This is particularly relevant to extant aquatic snakes who are deemed as having the most suitable physiology for scavenging.
A hypothesis put forth by \cite{sazima1990necrofagia} argued that chemical gradients in water would allow for a relatively easier detection of carrion.
This gained some support from \cite{devault2002scavenging}, who found a preponderence of aquatic snake species in their review of this behaviour.

@@

Interestingly, such comparisons have given insight into the feeding ecology of early hominins who, for instance, had the ability to craft tools for breaking open bones \citep{ARCM:ARCM12084}.
The question of where our ancestors placed on the hunter-scavenger axis during the Plio-Pleistocene has been a matter of debate for years.
A recent study investigating potential scavenging opportunities for hominins in Kenya found that, even when discounting bone material, there is a substantial amount of scavengeable meat left on predated remains; sufficient to sustain the requirements of an adult male \textit{Homo erectus} \citep{pobiner2015new}.
In some historical hominin-inhabited areas there were a greater number of felids than hyenids.
This is significant because hyenas are likely to have left far less flesh on a carcass than a felid such as a sabretooth enabling contemperaneous hominins to benefit \citep{pobiner2015new}.
The intelligence, resultant tool-use and cooperative nature of hominins meant they could likely adapt to take on more or less carrion depending on their environment \citep{moleon2014humans}.
%TG: or "could"? This is hypothetical behaviour stuff! Maybe past hyenas where exclusive sushi eaters!
%TG: Not sure about the intelligence thing: the sentence as the same meaning without and will be less anthropocentric sounding.

@@

\cite{shivik2006vultures} points out that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." 
But the fact remains that overcoming microorganism toxins is still a beneficial adaptation to any scavenger. 
Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
This adapation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans} so-called 'vultures of the seas') had an average pH of 1.5, which enables them to consume fisheries discards. 
Outside of the birds there is evidence of selection for 'toxification minimizers'.
From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
This implies later arrivers would benefit especially from well-developed detoxifying apparatus. 
\cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
Moreover, some authorities have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}.

%Being far less vagile than other vertebrate species,
%snakes are expected to develop detoxification strategies to overcome
%chemical defenses and make the best use of carrion.
%There is additional evidence that  Evolutionary pressures are not limited to competition
%for carcasses, of course, but evolutionarily, as snakes
%developed from eyeless fossorial species and radiated into terrestrial
%and arboreal predatory species (Rage 1994), an obligate
%scavenger evolutionary intermediate was likely.

%TG: Two points here: one speculation for the non-scavenging behaviour could be because of diet.
It is in the ability to process carrion that bats suffer.
Big bats (which are better suited for scavenging, following our previous argument) are typically frugivores and therefore lack the adaptations for digesting  meat.
While carnivorous bats are mainly found in the microbats which are insectivorous.
% Maybe they scavenge on dead insects but here we're talking about scavenging on real meat right?
%TG: Second point: bats fossil record is of course not so good but it's actually way better than birds! Most of the data we have is on teeth of course but it's probably enough for having a rough body mass and diet estimate! And that's much better than birds where you're stuck with just some fragments of long bones.

% Durophagy amount aquatic species, turtles, sharks, trigger fish, crocs etc.
